This report provides a comprehensive analysis of the YOLO (You Only Look Once) computer vision model, a state-of-the-art object detection algorithm that has significantly impacted various fields. The report traces the evolution of YOLO from its initial version to the latest iterations, highlighting the key architectural advancements and performance improvements that have occurred over time. It examines YOLO's core components, including backbone, neck, and head architectures, as well as the fundamental concepts of grid cells, bounding box prediction, confidence scores, and class probabilities. 

The report explores YOLO's diverse applications and substantial impact in biomedical science and agriculture through rigorous experimental evaluation. In biomedical applications, we assessed YOLO v11's performance on COVID-19 detection using chest X-rays (achieving mAP@50 of 0.726) and blood cell detection (reaching mAP@50 of 0.934). In agricultural applications, we evaluated fruit ripeness detection, where the model achieved exceptional accuracy (mAP@50 of 0.995). Our cross-domain comparison revealed that performance varies predictably based on task characteristics, with visually distinctive features yielding the best results.

The inherent strengths of YOLO, including real-time processing capabilities (33-200 FPS across applications), versatility across domains, and effectiveness with limited training data, are discussed alongside limitations such as challenges with subtle visual patterns and small overlapping objects. Novel domain-specific adaptations, including our Roughness-Local Binary Pattern (R-LBP) preprocessing technique for enhancing texture discrimination in fruit ripeness detection, demonstrate pathways to optimize performance for specialized applications.

The report concludes by outlining future research directions, including enhanced medical image processing, improved small object detection, architectural modifications incorporating attention mechanisms, computational efficiency optimizations, multi-modal integration, and cross-domain transfer learning. These advancements promise to further extend YOLO's utility across healthcare, agriculture, and numerous other fields requiring efficient and accurate object detection.